---
title: "autocorrelation"
author: "Collin"
date: "10/10/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(matrixcalc)
```

```{r}
#find pi function
find_pi = function(p){
  n = nrow(p)
  imp = diag(n) - t(p) #This is I - P^T
  
  imp[n,] = rep(1,n) #replace last row of I-P^T matrix with ones to make it invertible
  rhs = c(rep(0,n-1),1)
  
  return(solve(imp,rhs)) #
  
}
```



$\rho(k)$ is defined as the autocorrelation betwee $X_{i+k}$ and $X_i$. There is no dependency on `i` because the markov chain is *stationary*. In other words, $X_0 \sim \pi$, $X_1 \sim \pi$, ... $X_k \sim \pi$ $\forall \: i = 1,...k$, where `k` is the number of states in the markov chain. The distribution of $X_i$ does not depend on the $i^{th}$ state, so nor does the correlation.
```{r}
#function that computes the autocorrelation between X_i and X_{i+k}

MCautocor = function(k,P){
  #function has two parameters;
  #k: number of steps in the future
  #P: transition matrix of the markov chain 
  
  #Essentially need to find E(X_0 X_k). This will be the autocovariance, but then we can normalize to be between -1 and 1 by dividing by E(X_0 X_0) the second moment
  
  #first get the pis
  pi_vector = find_pi(P)
  m = nrow(P)
  
  #calculate the kth step transition matrix. Will need this for the probabilities in the next step (conditional exps)
  kth_step_P = matrix.power(P,k)

  
  
  #calculate conditional expectations (This will depend on k) E(X_k | X_0 = i) for all i in the state space i = 1,...m
  conditional_exp_vec = numeric(m)
  for(i in 1:nrow(P)){#loop over X_0 values (all the states 1,...,m)
    
    cond_exp = 0
    for(j in 1:nrow(kth_step_P)){ #calculate the conditional expectation for E(X_k | X_0 = # of the outer for loop)
      cond_exp = cond_exp + (j-1) * kth_step_P[j,i]  #j-1 for state 0, get rid of the minus 1 if state index starts at 1
    }
    #append this conditional expectation to the vector
    conditional_exp_vec[i] = cond_exp
    
  } 
  
  #sum the conditional expectations weighted by the pis from the pi vector
  #calculate E(X_k X_0): autocovariance
  acv = sum(pi_vector * conditional_exp_vec)
  return(c(acv, pi_vector, conditional_exp_vec))
  
}



```
`m` is not needed because it is already  known: `P` is an `m x m` matrix!


```{r}
#create functions that will be called in MCautocor_v2:
calc_EX = function(pi_vector, P_k){
  #this function will be called inside of MCautocor_v2, has two parameters:
  #pi_vector: which is exactly what it sounds like
  #P_k: the kth step transition matrix 
  m = nrow(P_k)
  EX_0 = 0
 
  for(i in 1:m){
    EX_0 = EX_0 + i*pi_vector[i]
    
  }
  return(EX_0)
}

calc_EXsquared = function(pi_vector, P_k){
  #this function will be called inside of MCautocor_v2, has two parameters:
  #pi_vector: which is exactly what it sounds like
  #P_k: the kth step transition matrix 
  m = nrow(P_k)
  EX_squared = 0
  
  for(i in 1:m){
    EX_squared = EX_squared + i^2 * pi_vector[i]
  }
  return(EX_squared)
}

pi_vec = c(1/3,1/3,1/3)
P <- matrix(c(1/3,1/3,1/3,
              1/3,1/3,1/3,
              1/3,1/3,1/3),nrow=3, ncol=3,byrow = TRUE)

calc_EX(pi_vec,P)
nrow(P)


```



#working version
```{r}
MCautocor_v2 = function(k,P){
  #function has two parameters
  #k: the kth step we are interested in
  #P: the transition matrix
  
  #get the pi vector
  pi_vector = find_pi(P)

  #get m
  m = nrow(P)
  
  #calculate the Kth step transition matrix
   kth_step_P = matrix.power(P,k)
  
 #calculate the necessary moments
    # E[X_0]
    EX_0 = calc_EX(pi_vector, kth_step_P)

    #E[X_0^2]
    EX_0_squared = calc_EXsquared(pi_vector, kth_step_P)

    #E[X_k X_0]
    conditional_exp_vec = numeric(m)
    for(i in 1:m){
      
      E_X_k_given_X_0_equals_i = 0
      #
      for(j in 1:m){
        E_X_k_given_X_0_equals_i = E_X_k_given_X_0_equals_i + j * kth_step_P[i,j]
        
      }
      conditional_exp_vec[i] = E_X_k_given_X_0_equals_i #append this E(X_k | X_0 = i) to the conditional_exp_vec
      
    }
    EX_kX_0 = sum(seq(1:m) *pi_vector*conditional_exp_vec)
    
    autocor = (EX_kX_0 - EX_0^2)/(EX_0_squared - EX_0^2)
    
    return(autocor)
}
  

```


```{r}
Ptest <- matrix(c(0.5, 0.25, 0.25, 0.75, 0.1, 0.15, 0.1, 0.2, 0.7), nrow = 3, byrow = TRUE)
MCautocor_v2(k=1, Ptest)
MCautocor_v2(k=2, Ptest)
MCautocor_v2(k=3, Ptest)
MCautocor_v2(k=4, Ptest)


```


```{r}
P <- matrix(c( 0.7, 0.2, 0.1, 
                 0.4, 0.6,   0, 
                 0,   1,   0  ), nrow=3, ncol=3,byrow = TRUE)
 kth_step_P = P
 k = 3
  for(i in 1:(k-1)){
    kth_step_P = kth_step_P %*% P #raising P to the kth power essentially
  }
 kth_step_P
```



