title: "die_game_markov_chain"
---
author: "Collin"
date: "10/6/2022"
output: html_document
---

```{r}
#Die Game: Example of Markov Chains
#One repeatedly rolls a die, keeping a running total. Each time the total exceeds 10, we receive one dollars, and continue play
#resuming where we left off, mod 10. ie, we have a total of 6, and we roll a 6. we receive a dollar, and start back at 2.
#Assume the player starts with 1 point

#code to obtain the transition matrix P. this game can be modeled with a discrete markov chain because the markov/memoryless property exists - > the future state (what you'll have after you roll) only depends on the current state (the total you currently have)
library(tidyverse)
library(LaplacesDemon) #for categorical distribtion
P = matrix(rep(0,100), nrow = 10) #initialize empty transition matrix
for(i in 1:10){ #loop over each row of the transition matrix
  
  #for the ith row, there are only 1-6 different possible outcomes from the dice roll
  for(j in 1:6){
    k = i + j
    if(k>10){
      k = k - 10
      P[i,k] = 1/6
    }
    
    
  }
  
}

```

```{r}
# is pi^T(I - P) = (I - P^T)pi ?
k = 3
pi_vector = matrix(c(.34,.43,.23))

Identity = diag(k)

P = matrix(c(.5,0,.5,
             .25,.5,.25,
             1,0,0), nrow = 3,byrow = TRUE)
Identity - P


#test for equality
t(pi_vector)%*%(Identity - P)
(Identity - t(P))%*%pi_vector

zero_vec = c(0,0,1)
A = (Identity - t(P))

A

solve(A, zero_vec)

A_with_O =  A
A_with_O[3,] = c(1,1,1) #replacing the last row of A with 1s (O vector)
new_zero_vector = c(0,0,1)
solve(A_with_O, new_zero_vector)

```



```{r}
#function to solve system of linear equations (markov chain) for pi

find_pi = function(p){
  n = nrow(p)
  imp = diag(n) - t(p) #This is I - P^T
  
  imp[n,] = rep(1,n) #replace last row of I-P^T matrix with ones to make it invertible
  rhs = c(rep(0,n-1),1)
  
  (solve(imp,rhs)) #
  
}



#use this one
alt_find_pi = function(p,k){
  niters = ceiling(log2(k))
  prd = p
  
  #calculate actual pi values
  actual_pis = find_pi(p)
  row1_error_vec = numeric(niters)
  row_average_error_vec = numeric(niters)
  
  for(i in 1:niters){
    prd = prd %*% prd
    
    #calculate the error at this iteration and append to a vector
    
    #error using row 1
    row1_error = sum(abs(prd[1,] - actual_pis))
    row1_error_vec[i] = row1_error #append
    
    
    #error averaging over the rows (colMeans)
    average_error = sum(abs(colMeans(prd) - actual_pis))
    row_average_error_vec[i] = average_error
  }
  return_list = list(actual_pis, row1_error_vec,row_average_error_vec)
  names(return_list) = c("actual_pis",
                         "row1_error_vector",
                         "row_average_error_vector")
  return(return_list)
  

}


```


#Find Pi (Continuous Markov Chain)
```{r}
find_pi_cont = function(q){
  n = nrow(q)
  q[n,] = rep(1,n)
  rhs = c(rep(0,n-1),1)
  pi_vec = solve(q,rhs) #solve for pi
  return(pi_vec)
}

```

# CTMC Example: Machine Repair
```{r}
Q = matrix(c(-.125, .05, 0,
             .25, -.175, .08,
             0, .125, -.08),
           byrow = TRUE, nrow = 3, ncol = 3)

find_pi_cont(Q)
```


```{r}
g = 4
for(i in 1:(g-1)){
  print(i)
}

```


```{r}
var_time = function(g){
  #g is the number of nodes on the network
  var_i_to_i_plus_1_vec = numeric(g-1)
  
  for(i in 1:(g-1)){
    #calculate the mean holding time at state i (This is the mean time to go from state i to i+1)
    var_i_to_i_plus_1 = 1/(i*(g-i))^2  #1/lambda^2
    var_i_to_i_plus_1_vec[i] = var_i_to_i_plus_1 #append to the vector,
    
  }
  
  return(sum(var_i_to_i_plus_1_vec))
  
  
}

#test
var_time(10)


```

```{r}


varTimeToAbsorb <- function(g){
  #calculate the mean time, will need it squared for final variance calculation
  expected_time_to_absorb = mean_time(g)
  
  #calculate the second moment...
  
  
  
}

print(varTimeToAbsorb(10))


```



```{r}

length(seq(-33,57,5))


```


```{r}
#3 heads in a row game: example
transition_matrix = matrix(rep(0,9), nrow = 3)
transition_matrix[1,1] = .5
transition_matrix[1,2] = .5
transition_matrix[2,1] = .5
transition_matrix[3,1] = 1
transition_matrix[2,3] = .5

transition_matrix
find_pi(transition_matrix)
```



```{r}
#Markov Chain Simulation: Hw2 problem 1
#create the transition matrix:

P = matrix(c(.5,.5,0,0,
             .5,0,.5,0,
             .5,0,0,.5,
             1,0,0,0), byrow = TRUE, nrow = 4)

P

#first, we can solve the system of linear equations directly: pi P = pi using the find_pi function defined previously
find_pi(P) 

#now lets see if we can reproduce these results by raising P to a very large power 
find_pi_power(P,k = 30)


P_squared = P%*%P

P_cubed = P_squared %*% P
pis = find_pi(P)

sum(abs(colMeans(P_cubed) - pis))



P_other_cubed = P %*% P %*% P

sum(abs(colMeans(P_other_cubed) - pis))



```

```{r}
#plot

```


```{r}
#Example using John's transition matrix 
#3x3

P <- matrix(c( 0.7, 0.2, 0.1, 
                 0.4, 0.6,   0, 
                 0,   1,   0  ), nrow=3, ncol=3,byrow = TRUE)

find_pi_power_out = alt_find_pi(P, k = 100)
find_pi_power_df = data.frame(n_iters = 1:length(find_pi_power_out$row_average_error_vector),
                              row1_error = find_pi_power_out$row1_error_vector,
                              row_average_error = find_pi_power_out$row_average_error_vector
                              ) %>% 
  pivot_longer(!n_iters,names_to = "error_type", values_to = "error")
  

#plot
ggplot(data = find_pi_power_df, mapping = aes(x = n_iters, y = error))+
  geom_smooth(mapping = aes(x = n_iters, y = error, color = error_type), se = FALSE)+
  scale_x_continuous(breaks = seq(1, length(find_pi_power_df$n_iters), by = 1))+
  ggtitle("Kth Step Transition Matrix Error - 3x3 Transition Matrix", subtitle = str_glue("K = {value}", value = max(find_pi_power_df$n_iters)))+
  xlab("# Iterations")+
  ylab("Error")+
    theme_minimal()+
  theme(plot.title = element_text(size = 16, face = "bold"))


#make this a function

```


```{r}
set.seed(2022)
#transition matrix generator
gen_matrix = function(dimension){
  #first create the empty matrix (all zeros)
  generated_matrix = matrix(0, nrow = dimension, ncol = dimension)
  
  
  #generate random numbers in a vector of length n, then normalize and assign to the ith row of the matrix
  for(i in 1:dimension){
    #generate the numbers
    rnum = sample(c(20,50,2345,3453,2323,1232),1)
    random_vec = runif(dimension, max = rnum)
    
    #normalize
    random_vec_normalized = random_vec/sum(random_vec)
   
    #assign to the ith row of the matrix
    generated_matrix[i,] = random_vec_normalized
  }
  return(generated_matrix)
}

transition_matrix_50 = gen_matrix(50) #create a 50x50 transition matrix

find_pi_power_out = alt_find_pi(transition_matrix_50,k = 100)
find_pi_power_df = data.frame(n_iters = 1:length(find_pi_power_out$row_average_error_vector),
                              row1_error = find_pi_power_out$row1_error_vector,
                              row_average_error = find_pi_power_out$row_average_error_vector
                              ) %>% 
  pivot_longer(!n_iters,names_to = "error_type", values_to = "error")
  

#plot
ggplot(data = find_pi_power_df, mapping = aes(x = n_iters, y = error))+
  geom_smooth(mapping = aes(x = n_iters, y = error, color = error_type), se = FALSE)+
  scale_x_continuous(breaks = seq(1, length(find_pi_power_df$n_iters), by = 1))+
  ggtitle(str_glue("Kth Step Transition Matrix Error - 50 x50 Transition Matrix"), subtitle = str_glue("K = {value}", value = max(find_pi_power_df$n_iters)))+
  xlab("# Iterations")+
  ylab("Error")+
    theme_minimal()+
  theme(plot.title = element_text(size = 16, face = "bold"))


find_pi_power_df


```


```{r}
#3 ALOHA model with 3 nodes, markov chain model
q = .3
p = .4

#row 1 of the transition matrix 9 (i=0)
p00 = (1-q)^3 + 3*q*p*(1-q)^2
p01 = 3*q*(1-q)^2*(1-p) + 6*q^2*p*(1-q)*(1-p)

p02 = 3*q^2*(1-q)*p^2 + 3*q^2*(1-q)*(1-p)^2 + 3*q^3*p*(1-p)^2

p03 = q^3*(1-p)^3 + q^3*p^3 + 3*q^3*p^2*(1-p)

row1 = c(p00,p01,p02,p03)


#row 2 of the transition matrix (i = 1)
p10 = p*(1-q)^2
p11 = (1-q)^2*(1-p) + 4*q*p*(1-q)*(1-p)
p12 = 2*q*(1-q)*(1-p)^2 + 2*q*(1-q)*p^2 + 3*q^2*p*(1-p)^2
p13 = q^2*(1-p)^3 + 3*q^2*p^2*(1-p) + q^2*p^3



row2 = c(p10,p11,p12,p13)


#row 3 of the transition matrix (i= 2)
p20 = 0
p21 = 2*(1-q)*p*(1-p)
p22 = 3*q*p*(1-p)^2 + (1-q)*(1-p)^2 + (1-q)*p^2
p23 = q*(1-p)^3 + 3*q*p^2*(1-p) + q*p^3



row3 = c(p20,p21,p22,p23)



#row 4 of the transition matrix (i = 3)
p30 = 0
p31 = 0
p32 = 3*p*(1-p)^2
p33 = (1-p)^3 + p^3 + 3*p^2*(1-p)
row4 = c(p30,p31, p32,p33)
sum(row4)

#putting it all together into the transition matrix
aloha_P = matrix(c(row1,
             row2,
             row3,
             row4), byrow = TRUE, nrow = 4,ncol = 4)
#find pi

aloha_pi = find_pi(aloha_P)

print(str_glue("Pi Vector: \npi_1 = {pi1} \npi_2 = {pi2} \npi_3 = {pi3}, \npi_4 = {pi4}", pi1 = aloha_pi[1],
               pi2 = aloha_pi[2],
               pi3 = aloha_pi[3],
               pi4 = aloha_pi[4]))


 #when q = .3, p = .6 about 18% of the epochs 0 active nodes
#about 24% of the epochs will have 1 active node
#about 30% of the epochs will have 2 active nodes
#about 28% will have 3 active nodes

#What is the long-run average time between collisions
#Use LOTP to get P(collision): P(C) = sum(i = 0 to 3) P(collision | state = i)
#Then do 1/P(Collision) to get average time (epochs) between collisions

p_coll_state0 = 3*q^2*(1-q)*p^2 + q^3*p^3 + 3*q^3*p^2*(1-p)
p_coll_state1 = 2*q*(1-q)*p^2 +  3*q^2*p^2*(1-p) + q^2*p^3
p_coll_state2 = (1-q)*p^2 + 3*q*p^2*(1-p) + q*p^3
p_coll_state3 = p^3 + 3*p^2*(1-p)
p_coll_state_i = c(p_coll_state0,p_coll_state1, p_coll_state2, p_coll_state3)

prob_collision = sum(aloha_pi*p_coll_state_i)

print(str_glue("Average time between collisions, assuming p = {p}, q = {q}: {time}",
               p = p, q = q, time = 1/prob_collision))

```


#Find Eta function
```{r}
#eta = E(T_in), average time to go from state i to state n
find_eta = function(transition_matrix){
  n = nrow(transition_matrix)
  q = diag(n) - transition_matrix
  
  q = q[1:(n-1), 1:(n-1)]
  ones = rep(1,n-1)
  solve(q,ones)
}


```



```{r}


expected_time = function(P){
  n = nrow(P) #dimensions of the transition matrix
  ones_vec = rep(1,n-1)
  
  ET_ij_matrix = matrix(0,n^2, nrow = n, ncol = n) #create the nxn matrix that will contain all of the expected times
  
  for(j in 1:n){ #loop over state space
    #for each iteration of this loop, I solve the system of linear equations (I - A)E = 1 for E
    
    #(I - A)
    imp = diag(n-1) - P[-j,-j] #I - A from my derivation
    
    
    #solving this system of linear equations will return an n-1 dimensional vector, containing
    #the expected times to get from state every state i to the jth state (the iteration of the for loop we're on)
    expectations_not_incl_j = solve(imp,ones_vec) #this is E(T_1j), E(T_2j), ... E(T_j-1,j), E(T_j+1,j)...E(T_nj)
    #now just need the last E(Tij), which is when i= j
    
    
    #this will give us the one expectation that we have to calculate directly
    E_ii = 1 + sum(P[j,-j]*expectations_not_incl_j) 
    
    #use append and after = j-1
    expectations_inc_j = append(expectations_not_incl_j,E_ii,after = j-1)
    
    #add this to the ET_ij_matrix, populate it with all the expected times to go from state i to state j
    ET_ij_matrix[,j] = expectations_inc_j
    
  }
  return(ET_ij_matrix)
  
}




generateTransitionMatrix <- function(numberRows){
  set.seed(1)
  transitionMatrix <- matrix(nrow=numberRows,ncol=numberRows)
  for (row in 1:numberRows){
    x <- runif(numberRows)
    rowValues <- x/sum(x)
    transitionMatrix[row,] <- rowValues
  }
  transitionMatrix
}

set.seed(1)
transitionMatrix <- generateTransitionMatrix(n)
expected_time(transitionMatrix)

```

```{r}


P[-1,1]
```
```{r}

P
```



```{r}
n = 3
P <- generateTransitionMatrix(n)
ones_vec = rep(1,n-1)
  
ET_ij_matrix = matrix(0,n^2, nrow = n, ncol = n)
  
for(j in 1:n){ #loop over state space
    imp = diag(n-1) - P[-j,-j] #I - A from my derivation
    
    expectations_not_incl_j = solve(imp,ones_vec) #this is E(T_1j), E(T_2j), ... E(T_j-1,j), E(T_j+1,j)...E(T_nj)
    #now just need the last E(Tij), which is when i= j
    
    E_ii = 1 + sum(P[-j,j]*expectations_not_incl_j)
    
    #use append and after = j-1
    expectations_inc_j = append(expectations_not_incl_j,E_ii,after = j-1)
    
    #add this to the ET_ij_matrix
    ET_ij_matrix[,j] = expectations_inc_j
}



#expectations_not_incl_j should be n-1 length
#expectations_incl_j should be length n



#test: try j = 2
j = 2
imp = diag(n-1) - P[-j,-j] #2x2 like it should be
P[-j,2]

expectations_not_incl_j = solve(imp,ones_vec)
expectations_not_incl_j #vector of length n-1, GOOD, this is the average time to get from state 1 to state 2, and state 3 to state 2. This does NOT include average time to get from state 2 to state 2
E_ii = 1 + sum(P[-j,j]*expectations_not_incl_j) #this will give us average time to get from state 2 to state 2
expectations_inc_j = append(expectations_not_incl_j,E_ii,after = j-1)
expectations_inc_j

ET_ij_matrix[,j] = expectations_inc_j
ET_ij_matrix


```


# Advice Nurse Problem: CTMC Simulation

## functions
### convert_tuple
```{r}
convert_tuple = function(N,k,m){
  integer_state = N*(m+1) + (k+1)

  return(integer_state)
}

convert_tuple(3,4,4)

```

### generate lambdas
```{r}
lambda_gen = function(isl){
  #isl -> integer states list
  # N is the number of nurses currently active
  #k is the number of people currently in the waiting room
  #n is total number of available nurses
  #m is the maximum capacity of the waiting room
  #mu is average call time so 1/mu is the rate at which a call ends
  # v is average intercall arrival time, so 1/v is the rate at which a new call arrives
  # p is the probability a new nurse is successfully added
  
  
  #other parameterization: function(integer_states_list) #integer_states_list is a vector of tuples (vectors) so the one
      #pass as argument here would be the jth element of integer_states_list, which will be a tuple, of which the first element
      #is N and the second element is k
  #N = integer_states_list[1]
  #k = integer_states_list[2]
  #n,m,mu,v will be variables coming from Q_gen(), so they'll be like global variables to lambda_gen.
  
  N = isl[1]
  k = isl[2]
 
  lambda_j = 0
  if(N == 0 & k < m){ #N = 0, k < m. no nurses and need the waiting room to fill up
    lambda_j = 1/v
    #print("a")
  }
  
  else if(N == 0 & k == m){ # N = 0, k = m. no nurses and the waiting room is full
    lambda_j = (1/v)*p
    #print("b")
  }
  
  else if((N >=1 & N <= n) & k < m){ # between 1 and n (inclusive ) nurses currently active, and waiting room is not full
    lambda_j = N/mu + 1/v
    #print("c")
      
  }
  
  else if((N >=1 & N < n) & k == m){ # 1 to n-1 nurses active, and waiting room is full
    lambda_j = N/mu + (1/v)*p
    #print("d")
  }
  
  #if(N == n & k < m){ #maximum number of nurses are active, waiting room is not full
   # lambda_j = n/mu + 1/v
   # print("e")
 # }
  
  if(N == n & k == m){ #maximum number of nurses are active and waiting room IS FULL
    lambda_j = n/mu
    #print("f")
    
  }
  return(lambda_j)
}

```

```{r}

```


```{r}


```

### generate transition probabilities
```{r}

integer_states_list = list(
  c(0,0), #1
  c(0,1), #2
  c(0,2), #3
  c(0,3), #4
  c(0,4), #5
  c(1,0), #6
  c(1,1), #7
  c(1,2), #8
  c(1,3), #9
  c(1,4), #10
  c(2,0), #11
  c(2,1), #12
  c(2,2), #13
  c(2,3), #14
  c(2,4), #15
  c(3,0), #16
  c(3,1), #17
  c(3,2), #18
  c(3,3), #19
  c(3,4) #20
)


mu = 2
v = 5
n = 3
m = 4
p = .9

#mu, n, v and p need to be defined globally for the function to work
lambda_C = 1/mu
lambda_I = 1/v 
tp_gen = function(jth_state, ith_state){
  #jth_state is the [[jth]] element of integer_states_list, and #ith_state is the [[ith]] element of integer_states_list
  
  #unpack the N's and K's
  jth_N = jth_state[1]
  jth_k = jth_state[2]
  ith_N = ith_state[1]
  ith_k = ith_state[2]
  
  p_ji = 0 #this is what we will be returning. This function is called inside of Q_gen
  
  if(jth_N == 0 & jth_k <= m){ #transition probabilities when 0 nurses are active
    if(ith_k == jth_k + 1 & ith_N == 0){ 
       p_ji = 1
      # print("ay yo")
    }else{
      p_ji = 0
    }
  }
  
  if((jth_N >=1 & jth_N <= n) & jth_k == 0){ #between 1 and n nurses active, 0 people in the waiting room
    
    if( jth_N == ith_N + 1 & jth_k == 0 & ith_k == 0){ # call ends and a nurse is removed
      p_ji = (jth_N*(lambda_C))/(jth_N*lambda_C + lambda_I)
      #print("Xd")
      
    }else if(jth_N == ith_N & ith_k == 1){ #call arrives
      p_ji = lambda_I / (jth_N*lambda_C + lambda_I)
    # print("HHFHDH")
    }else{
      p_ji = 0
     # print("here")
    }
    
  }
  
  if((jth_N >=1 & jth_N <= n) & jth_k < m & jth_k > 0){
    if(jth_N == ith_N & jth_k == ith_k + 1){ #call ends and waiting room isnt full
      p_ji = jth_N*lambda_C / (jth_N*lambda_C + lambda_I)
      #print("uwu")
      
    }else if(jth_N == ith_N & jth_k == ith_k - 1){ #call arrives and waiting room isn't full
      p_ji = lambda_I / (jth_N*lambda_C + lambda_I)
     # print("hm")
    }else{
      p_ji = 0
     # print("PLESE")
    }
  }
  
  if((jth_N >=0 & jth_N < n) & jth_k == m){ #waiting room full
    if (ith_N == jth_N + 1 & jth_k == m & ith_k == m){
      p_ji = lambda_I*p/(jth_N*lambda_C + lambda_I*p) #person arrives when the waiting room is full
     # print("wtf")
      
    }else if(jth_N == ith_N & jth_k == m & ith_k == m-1){
      p_ji = jth_N*lambda_C / (jth_N*lambda_C + lambda_I*p) #call ends and person leaves the waiting room when it was full
     # print("lmao")
    } else{
      p_ji = 0
    }
  }
  
  if( (jth_N == n & jth_k == m) & (ith_N == n & ith_k == m-1)){
    p_ji = 1 #the only way to go from state (n,m) to anywhere else is to n,m-1 and ithappens with probability 1
    #print("yup")
  }
  
  return(p_ji)
}

tp_gen(integer_states_list[[20]], integer_states_list[[3]])

```

### generate integer_state_list
```{r}
#another function that will be called inside of Q_gen
create_integer_state_list = function(n,m){
  integer_state = vector("list", (n+1)*(m+1))
iter = 1
for(N in 0:n){
  for(k in 0:m){
    integer_state[[iter]] = c(N,k)
    iter = iter + 1
  }
}
return(integer_state)

}


test_list = create_integer_state_list(3,4)
test_lambda_j = test_list[[2]][1]# second elemnt of this list is a vector (tuple state). need the first element
test_list
```



### test generate_lambda
```{r}

mu = 2
v = 5
lambda_gen(test_list[[4]]) #corresponding state (0,1). What is the rate you leave this state? -> .2

tp_gen(test_list[[4]],test_list[[5]])
```


### generate Q matrix
```{r}
Q_gen = function(n,m,p,mu,v){
  
  #define useful parameters:
  lambda_C = 1/mu
  lambda_I = 1/v
  dim = (n+1)*(m+1)
  
  Q = matrix(dim^2, nrow = dim, ncol = dim)
  
  integer_state_list = create_integer_state_list(n,m)
  
  for(i in 1:dim){
    for(j in 1:dim){
      lambda_j = lambda_gen(integer_state_list[[j]]) #the rate at which we leave state j
      p_ji = tp_gen(integer_state_list[[j]], integer_state_list[[i]])
      
      if(i == j){ #diagonals
        Q[i,j] = -lambda_j
        
      }else{
        Q[i,j] = lambda_j*p_ji
      }
    }
  }
  
 return(Q)
}


```


```{r}
test_Q = Q_gen(3,4,.9,2,5)
test_Q

```

```{r}

lambda_gen(test_list[[6]]) #corresponding state (0,1). What is the rate you leave this state? -> .2

tp_gen(test_list[[2]],test_list[[13]])
test_list
```

```{r}
find_pi_cont(test_Q)
test_pi = find_pi_cont(test_Q)
test_pi
```


## FUNCTIONS FOR CRITERIA OF INTEREST
### Proportion of Calls Rejected/Proportion of Time Nurses are Not Successfully Added
```{r}
prop_calls_rejected= function(pi_cont, n,m,p){
  #pi_cont is the vector of stationary probabilities for the CTMC
  #n is the number of nurses
  #m is the max number of the waiting room
  #p is the success probability of a nurse being successfully added to the system
  
  dim = (n+1)*(m+1)
  prop = 0
  for(j in 1:length(pi_cont)){
    if((j %% (m+1) == 0 & j != dim)){
      prop = prop + pi_cont[j]*(1-p)
      
    }else if(j == dim ){
      prop = prop + pi_cont[j]
      
    }else{
      next
    }
    
  }
  return(prop)

}
prop_calls_rejected(pi_cont = test_pi, n = 3,m=4,p=.9)


```

### Mean Number of Nurses Active (Patients talking to Nurse)
```{r}

mean_nurses = function(pi_cont,n,m){
  isl = create_integer_state_list(n,m) #will allow me to get N at each iteration 
  
  mean_number_nurses = 0
  for(j in 1:length(pi_cont)){
    #get value of N from jth state
    N = isl[[j]][1]
    mean_number_nurses = mean_number_nurses + pi_cont[j]*N
  }
  return(mean_number_nurses)
}

mean_nurses(test_pi,n = 3,m = 4)

```


### Mean Number of Patients in Waiting Room
```{r}
mean_patients_in_wr = function(pi_cont,n,m){
  isl = create_integer_state_list(n,m) #will allow me to get K at each iteration 
  
  mpwr = 0
  for(j in 1:length(pi_cont)){
    #get value of K from jth state
    K = isl[[j]][2]
    mpwr = mpwr + pi_cont[j]*K
  }
  return(mpwr)
}

mean_patients_in_wr(test_pi,n = 3,m = 4)

```


### Mean Number of Patients in the System
```{r}
#T = {total number of patients in the advice nurse system}
#E[T] = E(N) + E[K] , where N is number of active nurses (and effectively number of patients talking to nurses) and
#K is number of patients in the waiting room

mean_patients = function(pi_cont,n,m){
  mp = mean_patients_in_wr(pi_cont,n,m) + mean_nurses(pi_cont,n,m)
  return(mp)
}

mean_patients(test_pi,n = 3, m = 4)


```

### Mean Max Expected Waiting Time

This metric quantifies how long (on average) one must wait in the waiting room in the worst case scenario. ie, if you are in state (1,3), you have to wait on average 3/Nlambda_c.

It is slightly more complicated when you enter the waiting room when no nurses are currently active. ie, say you are in state (0,2). you need for the room to fill up ( at mean time 1/lambda_I) m-k times (so in our common example, this would be 4-2, or 2 more people), and then you would have to consider the mean time to go to (1,4) from (0,4) which is 1/(lambda_i*p), and ONLY THEN do you start taking into consideration the mean time for calls to start ending, which would be (k-1) times 1/lambda_C
```{r}
expected_max_waiting_time = function(pi_cont,n,m,v,mu,p){
  #get integer state list for single state to tuple state conversion
  isl = create_integer_state_list(n,m)
  mewt = 0
  
  for(j in 1:length(pi_cont)){
    N = isl[[j]][1]
    k = isl[[j]][2]
    
    if(N == 0 & k == 0){  #should be zero but already is so skip this iteration
      next
    }
    else if(N == 0 & k != 0){
      mewt = mewt + pi_cont[j]*(v*(m - k) + v/p + (k-1)*mu)
    }
    else{
      mewt = mewt + pi_cont[j]*(k*mu)/N
    }
    
  }
  return(mewt)
  
}

expected_max_waiting_time(test_pi,n = 3, m = 4, v = 5, mu = 2, p = .9)



```


### Mean Waiting Time
```{r}
mean_waiting_time = function(pi_cont,n,m,v){
  lambda = 1/v #rate of new arrivals into the system
  
  mp = mean_patients(pi_cont,n,m) #mean number of patients
  
  return(mp/lambda)
  
}
mean_waiting_time(test_pi, n = 3, m= 4, v = 5)

#SEe if this equals E(waiting room time) + E(call time)



```



```{r}
library(tidyverse)
c1 = c("1","2","3")
c2 = c("big","stupid","head")

out = crossing(c1,c2)
out

xy.list <- as.list(as.data.frame(t(out)))


```





### Score Calculator
```{r}

calculate_score = function(weights = c(1/4,1/4,1/4,1/4), mu, v, n, m, p){
  #vector of weights should be a 3 element vector with weights that sum to 1
  #weights should be in following order: (mean patients in waiting room, mean patients in system, mean number of nurses active)
  #default to (1/3,1/3, 1/3)
  
  #Cost = w1(mean_patients_in_waiting_room) + w2(prop_calls_rejected) + w3(mean_number_of_nurses_active)
  
  #mu is the mean call duration time
  #v is the mean intercall arrival time
  #n is number of nurses
  #m is max capacity of waiting room
  #p is success probability of a nurse being added when waiting room is full
  
  
  #get Q matrix
  Q = Q_gen(n = n, m = m, p = p, mu = mu, v = v)
  
  #get stationary probabilities (pi vector)
  pi_vector = find_pi_cont(Q)
  
  #get the score
  score = weights[1]*mean_patients_in_wr(pi_vector,n = n, m = m) + weights[2]*prop_calls_rejected(pi_vector, n, m,p) + weights[3]*mean_nurses(pi_vector, n,m) + weights[4]*expected_max_waiting_time(pi_vector,n = n, m = m, v = v, mu = mu, p = p)
  
  #should return a vector of the total score, and then each of the four components
  return(c(score,
           mean_patients_in_wr(pi_vector,n = n, m = m), #average number of patients in the waiting room
           prop_calls_rejected(pi_vector, n, m,p), # proportion of calls rejected
           mean_nurses(pi_vector, n,m), #average number of nurses active
           expected_max_waiting_time(pi_vector,n = n, m = m, v = v, mu = mu, p = p) #worst-case average waiting time
           ))
}

calculate_score(mu = 2, v = 5, n = n_test, m = m_test, p = p_test)



```



## Simulate Markov Chain function
```{r}
library(tidyverse)


n_test = c(3)
m_test = c(4)
p_test = c(.9)
simulate_MC = function(mu,v,n_vector, m_vector, p_vector){
  

  
  
  #create an empty dataframe that will store all of the scores and their respective parameterizations
  scores_df = data.frame(score = numeric(),interarrival_rate = numeric(), call_duration_rate = numeric(),
                         avg_pwr = numeric(),
                         prop_cr = numeric(),
                         avg_nurses_active = numeric(),
                         worst_case_mean_wait_time = numeric(),
                         n = numeric(), m =numeric(), p = numeric())
  
  
  
  #n_vector, m_vector, and p_vector are vectors containing all the different n,m, and p parameteriziations we want to ssearch over #essentially to see the behavior of the markov chain and see if there is a specific parameterization that minimizes our cost #function
  
  #listify n_vector, m_vector, p_vector into a list, where each element is a 3 element vector containing an n,m,p
  out = crossing(n_vector,m_vector,p_vector)
  parameter_list = as.list(as.data.frame(t(out)))
  
  
  #track the current minimum score
  curr_min_score = 0
  
  #loop over the values of the list
  for(i in 1:length(parameter_list)){
    #get the parameters for the current iteration
    n_i = parameter_list[[i]][1]
    m_i = parameter_list[[i]][2]
    p_i = parameter_list[[i]][3]
    
    #calculate the score for this iterations parameterization
    score_i = calculate_score(mu = mu, v = v, n = n_i, m = m_i, p = p_i)
    
    #append the score and its respective parameters to the dataframe that will contain all of them
    scores_df = scores_df %>% 
      add_row(score = score_i[1], interarrival_rate = 1/v, call_duration_rate = 1/mu, 
              avg_pwr = score_i[2],
              prop_cr = score_i[3],
              avg_nurses_active = score_i[4],
              worst_case_mean_wait_time = score_i[5],
              n = n_i, m = m_i, p = p_i)
    
    #check if this score is the new minimum, if so, update it
    if(score_i[1] < curr_min_score | i == 1){
      curr_min_score = score_i[1]
    }

}
    
    #create a new column for the scores dataframe, and have it be an indicator of whether or not that paramterization (row) is
    #the optimal one. Also normalize the score
  scores_df = scores_df %>% 
    mutate(minimizer = case_when(
      score == curr_min_score ~ "yes",
       TRUE ~ "no"
      )) %>% 
    mutate(score = score/sum(score))
  
  return(scores_df)
    
  
}

simulate_MC(mu = 2, v = 5, n_vector = n_test, m_vector = m_test, p_vector = p_test)

```


#How probability of successfully adding a nurse impacts key metrics
```{r}

#parameters
n_test = c(3)
m_test = c(4)
p_test = seq(.10,1,.01)

sim_out = simulate_MC(mu = 2, v = 5, n_vector = n_test, m_vector = m_test, p_vector = p_test)

#plot of worst-case mean wait time against p
ggplot(data = sim_out, mapping = aes(x = p, y = worst_case_mean_wait_time))+
  geom_smooth(mapping = aes(x = p, y = worst_case_mean_wait_time),se = FALSE)+
  ggtitle("Worst-Case Mean Wait Time & Proportion of Calls Rejected")+
  ylab("Time in Waiting Room (Worst Case)")+
  xlab("probability of sucessfully adding a nurse")+
  theme_minimal()





#plot of proportion of calls rejected against p
```





